---
title: "UNICEF DATA ANALYSIS"
subtitle: "RISK OF MATERNAL DEATH FOR TEENAGE GIRLS: GLOBAL PICTURE OF AN INEQUALITY"
author: "Angèle Etienne"
date: "04-20-2023"
format: html
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
install.packages("readr")
install.packages("ggplot2")
install.packages("ggplot")
install.packages("dplyr")
install.packages("cluster")
install.packages("MAP")
install.packages("mapdata")
install.packages("maps")
install.packages("rworldmap")
install.packages("rnaturalearthdata")
install.packages("rnaturalearth")
install.packages("sf")
library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
library(cluster)
library(MAP)
library(mapdata)
library(maps)
library(sf)
options(sf_use_s2 = FALSE)

unicef_indicator_2_1_<- read_csv("unicef_indicator_2 (1).csv")

```

***Every 2 minutes, a woman or a girl dies from complications related to pregnancy or childbirth (UNICEF)***

Every day in 2020, almost 800 women or girls died from preventable causes related to pregnancy and childbirth (World Health Organization). 

*This map highlights how the risk of maternal death is unevenly distributed across countries.*

```{r echo=FALSE}
options(sf_use_s2 = FALSE)
library(rnaturalearth)
library(sf)
library(ggplot2)

world <- ne_countries(scale = "medium", returnclass = "sf")
data <- left_join(world, unicef_indicator_2_1_, by = c("name" = "country"))

low_color <- "pink3" # Dark pink
middle_color <- "pink" # Light pink
high_color <- "#FFE4E1" # Very light pink

ggplot() +
  geom_sf(data = data, aes(fill = cut(obs_value, c(0, 500, 2000, Inf)))) +
  scale_fill_manual(values = c(low_color, middle_color, high_color), 
                    labels = c("0 to 500", "500 to 2000", "2000 to 65000"),
                    name = "Probability (1 in X)", drop = FALSE) +
  labs(title = "RISK OF MATERNAL DEATH FOR 15 YEAR-OLD GIRLS",
       subtitle = "Global overview (2020)",
       fill = "obs_value") +
  theme(plot.title = element_text(hjust = 0.5,family = "Bodoni", size = 12, color = "pink4"),
        plot.subtitle = element_text(hjust = 0.5, family = "Bodoni", size = 12, color = "pink4"),
        plot.margin = margin(1,1,1,1, "cm"))
```

Not all countries are equal when it comes to the risk of maternal mortality for teenage girls. 

*The disparities are glaring, as illustrated by the following chart.*


```{r echo=FALSE}
ggplot(unicef_indicator_2_1_, aes(x = obs_value, y = reorder(country, obs_value))) +
  geom_point(aes(color = obs_value), size = 2) +
  scale_color_gradient(low = "#FFE4E1", high = "pink3") +
  labs(title = "RISK OF MATERNAL DEATH FOR 15 YEAR-OLD GIRLS",
       subtitle = "Global overview (2020)",
       x = "Probability (1 in X)",
       y = "Countries") +
  theme(plot.title = element_text(hjust = 0.5, family = "Bodoni", size = 12, color = "pink4"),
        plot.subtitle = element_text(hjust = 0.5, family = "Bodoni", size = 12, color = "pink4"),
        plot.margin = margin(1,1,1,1, "cm"),
         axis.title.x = element_text(family = "Bodoni", size = 12, color = "pink4"),
        axis.title.y = element_text(family = "Bodoni", size = 12, color = "pink4"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text (family = "Bodoni", size = 12, color = "pink4"),
        panel.background = element_rect(fill = "#FAF3F0"))
```


Countries with similar risks can be identified.

*The following chart shows 3 different clusters with similar values in terms of risks.*

```{r include=FALSE}
X <- unicef_indicator_2_1_ $ obs_value
class(X)
kmeans <- kmeans(X, centers = 3)
kmeans$cluster
```

```{r echo=FALSE}

unicef_indicator_2_1_$color <- ifelse(unicef_indicator_2_1_$obs_value > 100, "pink","pink")
ggplot(unicef_indicator_2_1_,aes(x=factor(kmeans$cluster), y=obs_value, color=color))+
  geom_jitter(width = 0.2, height = 0, size = 3, alpha = 0.6) +
  labs(title = "RISK OF MATERNAL DEATH FOR 15 YEAR-OLD GIRLS (2020)",
       x = "Country clusters by kmeans",
       y = "Probability 1 in X") +
  theme(plot.title = element_text(hjust = 0.5, family = "Bodoni", size = 15, color = "pink4"),
        plot.subtitle = element_text(hjust = 0.5, family = "Bodoni", size = 12, color = "pink4"),
        plot.margin = margin(1,1,1,1, "cm"),
        panel.background = element_rect(fill = "#FAF3F0", color = NA),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(family = "Bodoni", size = 12, color = "black"),
        axis.title.y = element_text(family = "Bodoni", size = 12, color = "black"),
        axis.text.y = element_text(family = "Bodoni", size = 12, color = "pink4"),
        axis.text.x = element_text (family = "Bodoni", size = 12, color = "pink4"),
        legend.position = "none")
```

The inequalities between countries are even more apparent when looking at the extremes.

*The following chart shows that: the 5 countries with the lowest risk of maternal mortality for girls are located in Europe.*

*It is notable that between Belarus and Malta, 1st and 5th in the ranking, the risk is multiplied by more than 2.5.*

```{r echo=FALSE}

library(tidyverse)

  col_types = cols(
    country = col_character(),
    alpha_2_code = col_character(),
    alpha_3_code = col_character(),
    indicator = col_character(),
    sex = col_character(),
    unit_multiplier = col_integer(),
    obs_value = col_double(),
    numeric_code = col_character(),
    time_period = col_double(),
    time_period_activity_related_to_when_the_data_are_collected = col_logical(),
    unit_of_measure = col_character(),
    current_age = col_character(),
    observation_status = col_character(),
    observation_confidentaility = col_character() )
  
unicef_indicator_2_1_ %>%
  group_by(country) %>%
  summarise(obs_value = sum(obs_value, na.rm = TRUE)) %>%
  arrange(desc(obs_value)) %>%
  head(5) %>%
  ggplot(aes(x = fct_reorder(country, -obs_value), y = obs_value)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "COUNTRIES WITH THE LOWEST RISK OF MATERNAL DEATH  FOR
15 YEAR-OLD GIRLS (2020)",
       x = "Country",
       y = "Probability 1 in X") +
  theme(plot.title = element_text(hjust = 0.5, family = "Bodoni", size = 15, color = "pink4"),
           axis.title.x = element_text(family = "Bodoni", size = 12, color = "black"),
        axis.title.y = element_text(family = "Bodoni", size = 12, color = "black"),
        axis.text.y = element_text(family = "Bodoni", size = 8, color = "pink4"),
        axis.text.x = element_text (family = "Bodoni", size = 6, color = "pink4"),
        plot.margin = margin(1,1,1,1, "cm"))
```

*The following bar chart shows the 5 countries with the highest risks of maternal death for 15 year-old-girls. The countries are located in sub-Saharan Africa.*


```{r echo=FALSE}
library(tidyverse)

  col_types = cols(
    country = col_character(),
    alpha_2_code = col_character(),
    alpha_3_code = col_character(),
    indicator = col_character(),
    sex = col_character(),
    unit_multiplier = col_integer(),
    obs_value = col_double(),
    numeric_code = col_character(),
    time_period = col_double(),
    time_period_activity_related_to_when_the_data_are_collected = col_logical(),
    unit_of_measure = col_character(),
    current_age = col_character(),
    observation_status = col_character(),
    observation_confidentaility = col_character() )

unicef_indicator_2_1_ %>%
  group_by(country) %>%
  summarise(obs_value = sum(obs_value, na.rm = TRUE)) %>%
  arrange(obs_value) %>%
  head(5) %>%
  ggplot(aes(x = fct_reorder(country, obs_value), y = obs_value)) +
  geom_bar(stat = "identity", fill = "pink3") +
  labs( title = "COUNTRIES WITH THE HIGHEST RISK OF MATERNAL DEATH FOR 15 YEAR-OLD GIRLS (2020)",
    x = "Country",
    y = "Probability 1 in X" ) +
  theme(
    plot.title = element_text(hjust = 0.5, family = "Bodoni", size = 15, color = "pink4"),
    axis.title.x = element_text(family = "Bodoni", size = 12, color = "black"),
    axis.title.y = element_text(family = "Bodoni", size = 12, color = "black"),
    axis.text.y = element_text(family = "Bodoni", size = 8, color = "pink4"),
    axis.text.x = element_text(family = "Bodoni", size = 6, color = "pink4"),
    plot.margin = margin(1, 1, 1, 1, "cm")
  )
```

**In Chad, a 15-year-old girl has a 1 in 15 chance of dying of maternal causes. She is therefore 4,000 times more likely to die from maternal causes than a 15-year-old girl in Belarus.** 

